---
title: "Linear_regression"
author: "Fernando Loaiza"
date: "`r Sys.Date()`"
output: html_document
---

https://github.com/sajal2692/data-science-portfolio?tab=readme-ov-file

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The data used in this analysis comes from *Principles of Econometrics* by Hill, Griffiths, and Lim (Wiley, 2008). The dataset `br.csv` includes 1080 observations on the prices and characteristics of houses sold in Baton Rouge, LA in 2005. The variables considered in the analysis are:

- **price**: Price of the house (in dollars)
- **sqft**: Total square feet
- **bedrooms**: Number of bedrooms
- **baths**: Number of full baths
- **age**: Age of the house in years
- **Occupancy**: Owner = 1, Vacant = 2, Tenant = 3
- **Pool**: Yes = 1, No = 0
- **Style**: 
  - Traditional = 1
  - Townhouse = 2
  - Ranch = 3
  - New Orleans = 4
  - Mobile Home = 5
  - Garden = 6
  - French = 7
  - Cottage = 8
  - Contemporary = 9
  - Colonial = 10
  - Acadian = 11
- **Fireplace**: Yes = 1, No = 0
- **Waterfront**: Yes = 1, No = 0
- **DOM**: Days on the market

## Data Loading and Preliminary Analysis

```{r}
# Load the data
br <- read.table("/Users/fernandoloaizae/Library/CloudStorage/OneDrive-Personal/GitHub/Data_Projects/Data_Science/lab1/br.csv", sep = ",", header = TRUE)
attach(br)

# Histogram of price
hist(price, 30, col = "blue", main = "Histogram of Price")

# Histogram of log-transformed price
hist(log(price), 30, col = "red", main = "Histogram of Log-Transformed Price")

# Scatterplot of prices vs square feet
plot(sqft, price, col = "red", main = "Price vs Square Feet")

# Scatterplot of prices vs age
plot(Age, price, col = "red", main = "Price vs Age")

# Boxplot of price by Occupancy
plot(factor(Occupancy), price, col = 'red', main = "Price by Occupancy")

# Boxplot of price by Style
plot(factor(Style), price, col = 'yellow', main = "Price by Style")

# Pair plot of price against square feet and age
pairs(price ~ sqft + Age, upper.panel = panel.smooth)


# Linear regression model
regr <- lm(price ~ sqft + Age + Pool + Bedrooms + Fireplace + Waterfront + DOM)
summary(regr)

# OLS coefficients
coefficients(regr)

# Covariance matrix for model parameters
CovBeta <- vcov(regr)
CovBeta[2,2]  # Variance of beta_1_hat
sqrt(CovBeta[2,2])  # Standard error of beta_1_hat

# t statistic for beta_1_hat
coefficients(regr)[2] / sqrt(CovBeta[2,2])

# Confidence intervals for model parameters
confint(regr, level = 0.95)

# Fitted values and residuals
yf <- fitted(regr)
e <- residuals(regr)

# Property of the residuals: zero mean
mean(e)

# Histogram of residuals
hist(e, 60, col = 'blue', main = 'Histogram of Residuals')

# Correlation checks
cor(yf, e)  # Residuals uncorrelated with fitted values
cor(sqft, e)  # Residuals uncorrelated with sqft

# Plot of predicted vs observed values
plot(price, yf, col = "red", main = "Predicted vs Observed Prices")

# Plot of residuals vs observed values
plot(price, e, col = "red", main = "Residuals vs Observed Prices")
abline(h = 0)

# Plot of residuals vs predicted values
plot(yf, e, col = "red", main = "Residuals vs Predicted Prices")
abline(h = 0)

# Hat values (leverages)
h <- hatvalues(regr)
sum(h)  # Trace(H) = p+1
mean(h)

# Barplot of hat values
barplot(h, col = 'blue', main = "Hat Values")
abline(h = mean(h), col = 'red')

# Standardized residuals
r <- rstandard(regr)

# Plot of standardized residuals vs fitted values
plot(yf, r, main = "Standardized Residuals vs Fitted Values")

# Plot of standardized residuals vs hat values
plot(h, r, main = "Standardized Residuals vs Hat Values")

# Regression model with categorical variables
regr2 <- lm(price ~ sqft + Age + Pool + Bedrooms + Pool + Fireplace + Waterfront + DOM + factor(Occupancy) + factor(Style))
summary(regr2)

# Model updates by adding and removing variables
summary(update(regr2,  .~. - Fireplace))
summary(update(regr2,  .~. - factor(Occupancy)))
```


